How to make an abridged life table and calculate life expectancy

### 2016-09-21 working_teaching_examples/how_to_abridged_life_table.Rmd
added cod decomp examples

head(canada.men)
head(quebec.men)
```

@@

@@

first.strata.cod <- alabama.by.strata.cod[alabama.by.strata.cod$StateYearRaceSex == "Alabama.2013.White.Male", ]
second.strata.cod <- alabama.by.strata.cod[alabama.by.strata.cod$StateYearRaceSex == "Alabama.2013.Black.Male", ]

@@

cod.tab.alabama.male.1969$prop1 <- second.strata.cod$cod_prop_death5

@@

system.time(first.strata.life.table <- data.frame(life.table(data = first.strata, age.groups = "Age2", num.ages.in.group = "nx", death.counts = "total_deaths5", population.counts = "Population")))

second.strata.life.table <- data.frame(life.table(data = second.strata, age.groups = "Age2", num.ages.in.group = "nx", death.counts = "total_deaths5", population.counts = "Population"))

first.strata.life.table[, c("Age2", "total_deaths5", "Population", "R_x", "q_x", "l_x", "d_x", "L_x", "T_x", "e_x")]

system.time(
for(stratum.id in levels(droplevels(alabama.by.strata$StateYearRaceSex))) {
  assign(paste0("lt.", as.character(stratum.id)), 
         life.table(data = subset(alabama.by.strata, StateYearRaceSex == stratum.id), 
                                  age.groups = "Age2", num.ages.in.group = "nx", 
                                  death.counts = "total_deaths5", population.counts = "Population"))
}
)
```

>>>>

head(canada.men)
head(quebec.men)

cod.props <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteGap/working_teaching_examples/auger_cod_proportions.csv")

@@

```{r calc_cod_decomp}
canada.cod.decomp <- cause_of_death_decomp(life.table1 = life.table.qm, life.table2 = life.table.cm, decomp.table = age_decomp,
                      cod.table = cod.props,age.colname.cod.table = "Age", COD.colname.cod.table = "cod", 
                      prop1.colname.cod.table = "prop1", prop2.colname.cod.table = "prop2")

cod.contribution$Cause.of.death <- factor(cod.contribution$Cause.of.death, 
                                          levels = cod.contribution$Cause.of.death[order(cod.contribution$cod.contribution,
                                                                                         decreasing = T)])

ggplot(cod.contribution, aes(x = Cause.of.death, y = cod.contribution)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_abline(intercept = 0, slope = 0)
```

@@

@@

cod.tab.alabama.male.1969$prop1 <- second.strata.cod$cod_prop_death5
cod.tab.alabama.male.1969 <- cod.tab.alabama.male.1969 %>% arrange(COD2, Age2)

@@

system.time(life.table.whites <- data.frame(life.table(data = first.strata, age.groups = "Age2", num.ages.in.group = "nx", death.counts = "total_deaths5", population.counts = "Population")))

life.table.blacks <- data.frame(life.table(data = second.strata, age.groups = "Age2", num.ages.in.group = "nx", death.counts = "total_deaths5", population.counts = "Population"))

life.table.whites[, c("Age2", "total_deaths5", "Population", "R_x", "q_x", "l_x", "d_x", "L_x", "T_x", "e_x")]

system.time(
for(stratum.id in levels(droplevels(alabama.by.strata$StateYearRaceSex))) {
  assign(paste0("lt.", as.character(stratum.id)), 
         life.table(data = subset(alabama.by.strata, StateYearRaceSex == stratum.id), 
                                  age.groups = "Age2", num.ages.in.group = "nx", 
                                  death.counts = "total_deaths5", population.counts = "Population"))
}
)

alabama_age_decomp <- le_age_decomp(life.table.blacks, life.table.whites, "Age2")
alabama_age_decomp

ggplot(alabama_age_decomp, aes(y = C_x, x = Ages)) + geom_point() + scale_y_continuous("Contribution of Age to Gap")
sum(alabama_age_decomp$C_x)
life.table.whites$e_x[1] - life.table.blacks$e_x[1]

ggplot(alabama_age_decomp, aes(y = Life_Expectancy_1, x = Ages)) + geom_point(aes(col="Black")) + scale_y_continuous("Life Expectancy (Yrs)") +
  geom_point(aes(y = Life_Expectancy_2, col="White"))
```
